Detection of Oscillations in Aldebaran with Ground-Based Observations

### 2017-09-22 paper/aldebaran.tex
Added Seismic text

@@

\author{Guy R. Davies}

@@

@@

@@

In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time. We validate this method and its result with new RV observations with the SONG~Telescope, and photometry from the \ktwo Mission. By measuring the average frequency \numax of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.17 \pm 0.05$~\msun. This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that tehy may have once been habitable.

@@

@@

\begin{figure}
\centering
% \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
\includegraphics[width=0.75\textwidth]{Aldebaran_lc.eps}
\caption{\ktwo lightcurve of Aldebaran.}
\label{k2_lightcurve}
\end{figure}

@@

@@

>>>>

\newcommand{\guy}[1]{{\bf \color{blue} #1}}

@@

\author[0000-0002-4290-7351]{Guy R. Davies}

\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}
\affiliation{Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}

@@

\author{Tim White}
\affiliation{Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}

@@

\author{Andrea Miglio}

\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}
\affiliation{Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}

\author{Mikkel N. Lund}

\affiliation{Stellar Astrophysics Centre, Department of Physics 
and Astronomy, Aarhus University, Ny Munkegade 120, DK-8000 Aarhus C, Denmark}
\affiliation{School of Physics and Astronomy, University of
  Birmingham, Birmingham, B15 2TT, United Kingdom}

@@

In this paper, we present a re-analysis of these original RV data in which we not only confirm this signal, but detect acoustic oscillations in Aldebaran for the first time. We validate this method and its result with new RV observations with the SONG~Telescope, and photometry from the \ktwo Mission. By measuring the frequency of maximum power ,\numax, of these \emph{p}-mode oscillations we asteroseismically determine the mass of Aldebaran to be $1.16 \pm 0.07$~\msun. This measured stellar mass allows us to calculate that any moons of this giant planet, although they are now likely to be very hot, may have had equilibrium temperatures comparable to that of the Earth when Aldebaran was on the main sequence, raising the possibility that they may have once been habitable.

@@

\section{Asteroseismology of Red Giants}
Asteroseismology is a powerful tool for the charaterisation of red giant stars \citep[see][for a detailed review]{2013ARA&A..51..353C}.  Red giants are solar-type pulsators, they exhibit acoustic modes of oscillation that are both excited and damped by stellar convection.  In the frequency-power spectrum of either radial velocity of photometric observations, these modes of oscillation form an often intricate pattern of sharp peaks where the properties of the pattern can be mapped back to intinsic stellar properties \citep[e.g.,][]{2016AN....337..774D}.  The easiest properties of the pattern to determine are the the frequency of maximum amplitude \numax and the so-called large separation \Dnu \citep{1995A&A...293...87K}.  The two observables are often referred to as the global average seismic parameters.

Stellar properties can be determined with the average seismic measures and effective temperature \teff.  A set of scaling relations exist that illustrate the constraint the global asteroseismic measures provide.  These scaling relations are understood to be approximate but never the less are commonly used for radius $R$, mass $M$, and surface gravity $g$:
\begin{eqnarray}
\left( \frac{R}{\mathrm{R_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right) \, 
\left( \frac{\Delta \nu}{\Delta \nu_{\odot}} \right)^{-2} \, \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}, \\
\left( \frac{M}{\mathrm{M_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right)^{3} \, 
\left( \frac{\Delta \nu}{\Delta \nu_{\odot}} \right)^{-4} \, \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{1.5}, \\
\left( \frac{g}{\mathrm{g_{\odot}}} \right) \simeq & \left( \frac{\nu_{\rm max}}{\nu_{\rm max, \odot}} \right) \, 
 \left( \frac{T_{\rm eff}}{T_{\rm eff, \odot}} \right)^{0.5}. \\
\end{eqnarray}
While the accuracy of these scaling relations is still a matter of ongoing work \citep[e.g.,][]{2017ApJ...844..102H}, stellar parameters can be estimated by comparing observables to parmaeter from models of stellar evolution.

As can be seen from the scaling relations, asteroseismology can provide constraint on Mass, Radius, and surface gravity if \Dnu, \numax, and \teff can be measured.  In practise, for the higher luminosity red giants it is more straight forward to measure \numax than \Dnu.  Typical values for \numax might range from $0.1$ to $20 \rm \mu Hz$ for the luminous giants and $20$ to $50 \rm \mu Hz$ for red giants with luminosities similar to the red clump .  For the same sets of stars, \Dnu would take ranges from $0.02$ to $3 \rm \mu Hz$ and $3$ to $7 \rm \mu Hz$ depending on stellar mass and \teff (or metallicity which is considered more fundamental) \citep[e.g.,][]{2011A&A...525L...9M, 2013A&A...559A.137M}.  Hence, \numax requires less frequency resolution than \Dnu to establish a good measurement.

The ability to measure \numax and \Dnu, therefore, depends on the length and sampling rate of a data set.  For regularly spaced data, typically analysed using an estimate of the frequency-power spectrum, the Nyquist limit and frequency resolution are trivial to estimate.  For the historic data sets proposed here the typical length is \guy{blah} and typical sampling is \guy{blah}.  This in principle gives us access to measures of \numax for \guy{these types of stars} and \Dnu $\,$ \guy{for those types of stars}.

@@

\begin{figure}
\centering
% \includegraphics[width=0.75\textwidth]{Aldebaran_lc.png}
\includegraphics[width=0.75\textwidth]{aldebaran_lc.eps}
\caption{\ktwo lightcurve of Aldebaran.}
\label{k2_lightcurve}
\end{figure}

@@

We have determined the stellar properties of Aldebaran with the additional constraint of \numax determined from our analysis here.  The details of the stellar modelling are presented in detail in \ref{sec:mod}.  We have considered the impact of our additional \numax constraint and run our analysis for different sets of spectroscopic estimates.  For final values we adopt the Sheffield et al. (2012) b spectrosocpic solution as being 'middle of the road' estimates.  For the analysis without constraint on \numax we find estimates of the stellar propeties as $M = 1.43^{+0.24}_{-0.2} \, \mathrm{M_{\odot}}$ and age $4.86^{3.56}_{2.04} \, \rm Gyr$.  With the inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and age $6.38^{1.42}_{1.12} \, \rm Gyr$.  It is clear that the addition of \numax provides substatially more precise estimates of Mass and Age.

@@

\include{aldeb_North}
